### 2022-04-05 fivebranes/boundary.tex
boundary stuff

\subsection{Defects in Holomorphic-Topological Field Theories}

\parsec

Consider a $\Z/2$-graded holomorphic-topological field theory on a split THF manifold $X\times M$ where $X$ is a complex manifold and $M$ is a smooth manifold, and let $\cL$ denote the underlying holomorphic-topological local Lie algebra. Suppose in addition, we have a  defect supported along the image of a closed embedding \[\iota : N\subset X\times M.\]

\parsec

Following \cite{PW}, the algebra of local operators on the defect is naturally a quotient of the koszul dual

\parsec

We give an operational description of the algebra of local operators on the defect. The description is meant to codify the idea that the algebra can be computed as an algebra of boundary observables for the theory compactified on the link of the defect.

Consider the normal bundle $T_{(X\times M)/N}$ and choose a fiberwise metric $h$. There is a natural surjection

\[r = (\pi, \|-\|^{2}): T_{(X\times M)/N}\to N \times \R_{\geq 0}\]

where the first component is the natural projection, and the second component is the norm squared with respect to $h$. This equips the normal bundle with the structure of a stratified vector bundle over $N\times \R_{\geq 0}$.

We wish to consider the pushforward $r_{*}\cL$. Note that away from the zero section, the restriction of $r$ gives us the structure of a sphere bundle over $N\times \R_{>0}$. Moreover, the fiber spheres are equipped with canonical foliations given by

\[\]

Let's describe its sections explicitly.

\begin{itemize}
  \item For an open set
  \item
\end{itemize}