### 2022-06-07 fivebranes/membrane.tex
merge

\documentclass[11pt]{amsart}

%\usepackage{../macros-master}
\usepackage{macros-fivebrane}

\begin{document}

\section{membrane}

In this section we provide a description of the holomorphic, or minimal, twist of the worldvolume theory on a single membrane. 
It is defined on any three-manifold of the form $L \times \Sigma$ where $L$ is a real oriented one-manifold and $\Sigma$ is a Riemann surface equipped with a fourth root of the canonical bundle $K^{1/4}_\Sigma$.

The theory is $\Z/2$ graded which combines the cohomological $\Z$ and parity $\Z/2$ gradings present in the untwisted theory.
In the BV formalism the fields of the theory are
\[
\Omega^\bu(L) \otimes \Omega^{0,\bu}(\Sigma, K^{1/4}_\Sigma \otimes \C^4 \oplus \Pi K^{3/4} \otimes \C^4) .
\]
We denote fields according to the direct sum above by pairs $(\phi_a,\chi_a)$ where $a=1,\ldots 4$. 
The full BV action is 
\[
\int_{L \times \Sigma} \ep^{ab} \phi_a \d_{dR} \chi_b + \int_{L \times \Sigma} \ep^{ab} \phi_a \dbar \chi_b .
\]
The equations of motion are simply 
\[
\d_{dR} \phi_a + \dbar \phi_a = \d_{dR} \chi_a + \dbar \chi_a = 0
\]
for $a=1,\ldots, 4$.

In the (degenerate) BV formalism, this free theory can be packaged into the data of the abelian $\Z/2$ graded local Lie algebra on $L \times \Sigma$ 
\[
\cL_{membrane} = \Pi \Omega^\bu(L) \otimes \Omega^{0,\bu}(\Sigma, K^{1/4}_\Sigma \otimes \C^4 \oplus K^{3/4} \otimes \C^4)
\]
where the $\Pi$ denoted parity shift.

\section{$E(1|6)$}

\end{document}