### 2022-08-07 fivebranes/main.tex
main

\documentclass[11pt]{amsart}

%\usepackage{../macros-master}
\usepackage{macros-fivebrane}

%\linespread{1.2} %for editing

\addbibresource{references.bib}

\begin{document}

\title{A holomorphic approach to branes in twisted~$M$-theory}
\maketitle

\begin{abstract}
abstract here
\end{abstract}

\tableofcontents

\input{intro}

%Twisted supergravity
\input{twisted}

%The character of holomorphic local operators
%%introduce local operators for hol/top theories.
%%introduce hol characters for hol/top theories. 
%%example of a single fivebrane
%%example of a single membrane
%%Large N limit.
\input{index}

%Factorization algebras in twisted M theory
%%`Restricted' factorization algebra.
%%Koszul duality.

\input{fact}
%A holographic decomposition of local operators
%%E(3|6) decomposition of local operators. 
%%character computations following KR resolutions.
%%E(1|6) decomposition of local operators. 
%%some finite N membrane indices??

\input{ops}

\input{kr}

%%backreaction for single fivebrane.

%%\input{??}

\end{document}