### 2022-10-13 fivebranes/main.tex
abstract

\documentclass[11pt]{amsart}

%\usepackage{../macros-master}
\usepackage{macros-fivebrane}

\linespread{1.15} %for editing

\addbibresource{references.bib}

\begin{document}

%\title{A holomorphic approach to fivebranes}
\title{A holographic approach to the six-dimensional superconformal index}
\maketitle

\begin{abstract}
We present a conjectural description of the algebra of operators on a stack of finitely many fivebranes in $M$-theory at the level of the holomorphic twist.
Our approach is through the lens of twisted holography and utilizes a description of the minimal twist of eleven-dimensional supergravity. 
Strikingly, we find that local operators are modules for the exceptional linearly compact super Lie algebra $E(3|6)$. From the conjectural description of local operators we deduce closed formulas for the superconformal index of six-dimensional $\cN=(2,0)$ theories of type $A_{N-1}$.
\end{abstract}

\setcounter{tocdepth}{1}
\tableofcontents

\input{intro}

\subsection*{Acknowledgements}
We thank K.~Costello, D.~Gaiotto, N.~Garner, F.~Hahner, Si Li, N.~Paquette, and P.~Yoo for conversations and discussions about ideas that helped lead to the completion of this work. 
Special thanks are due to Ingmar Saberi for his collaborations on past, current, and future related projects.
The work of S.R. is supported by the Perimeter Institute for Theoretical Physics. Research at Perimeter Institute is supported in part by the Government of Canada, through the Department of Innovation, Science and Economic Development Canada, and by the Province of Ontario, through the Ministry of Colleges and Universities.
The work of B.W. was supported by Boston University and the University of Edinburgh. 
%Twisted supergravity
\input{twisted}

%The character of holomorphic local operators
%%introduce local operators for hol/top theories.
%%introduce hol characters for hol/top theories. 
%%example of a single fivebrane
%%example of a single membrane
%%Large N limit.
\input{index}

%Factorization algebras in twisted M theory
%%`Restricted' factorization algebra.
%%Koszul duality.

\input{fact}
%A holographic decomposition of local operators
%%E(3|6) decomposition of local operators. 
%%character computations following KR resolutions.
%%E(1|6) decomposition of local operators. 
%%some finite N membrane indices??

\input{ops}

\input{finite}

%%backreaction for single fivebrane.

%%\input{??}

\printbibliography

\end{document}