### 2022-12-27 aone/main.tex
elliptic

\documentclass{amsart}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage[titletoc]{appendix}
\usepackage{hyperref}

%\usepackage{tikz} 	  % Replaced with tikz-cd, but am
\usepackage{tikz-cd}  % willing to learn tikz if necessary
	\usetikzlibrary{decorations.pathmorphing}

\usepackage[all]{xy}
\usepackage{stmaryrd}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{tabularx}

\usepackage{macros}

\renewcommand\tabularxcolumn[1]{m{#1}}
\newcolumntype{Y}{>{\centering\arraybackslash}X}

\begin{document}

\title{A Pair of Holomorphic Fivebranes}

\date{\today}
\maketitle

%background

\section{A Single Fivebrane}

% descent of brane scan cocycle on \widehat{E(5|10)} = local anomaly polynomial from backreaction

% generators and relations for mode algebras on different 5-manifolds

% central charge matching

\section{Two Fivebranes}
% central extension of E(3|6)

% codimension 2 defects and the two different localizaitons

% dim red and comparison with 5d N=2 + oper type boundary condition

\section{Coproducts}

\end{document}