### 2023-02-07 fivebranes/membrane.tex

\documentclass[11pt]{amsart}

%\usepackage{../macros-master}
\usepackage{macros-fivebrane}

\begin{document}

\section{membrane}

In this section we provide a description of the holomorphic, or minimal, twist of the worldvolume theory on a single membrane. 
It is defined on any three-manifold of the form $L \times \Sigma$ where $L$ is a real oriented one-manifold and $\Sigma$ is a Riemann surface equipped with a fourth root of the canonical bundle $K^{1/4}_\Sigma$.

The theory is $\Z/2$ graded which combines the cohomological $\Z$ and parity $\Z/2$ gradings present in the untwisted theory.
In the BV formalism the fields of the theory are
\[
\Omega^\bu(L) \otimes \Omega^{0,\bu}(\Sigma, K^{1/4}_\Sigma \otimes \C^4 \oplus \Pi K^{3/4} \otimes \C^4) .
\]
We denote fields according to the direct sum above by pairs $(\phi_a,\chi_a)$ where $a=1,\ldots 4$. 
The full BV action is 
\[
\int_{L \times \Sigma} \ep^{ab} \phi_a \d_{dR} \chi_b + \int_{L \times \Sigma} \ep^{ab} \phi_a \dbar \chi_b .
\]
The equations of motion are simply 
\[
\d_{dR} \phi_a + \dbar \phi_a = \d_{dR} \chi_a + \dbar \chi_a = 0
\]
for $a=1,\ldots, 4$.

In the (degenerate) BV formalism, this free theory can be packaged into the data of the abelian $\Z/2$ graded local Lie algebra on $L \times \Sigma$ 
\[
\cL_{membrane} = \Pi \Omega^\bu(L) \otimes \Omega^{0,\bu}(\Sigma, K^{1/4}_\Sigma \otimes \C^4 \oplus K^{3/4} \otimes \C^4)
\]
where the $\Pi$ denoted parity shift.

\section{$E(1|6)$}

\subsection{Operators on a single membrane}

\begin{prop}
\label{prop:3done}
The local character $\chi_{\lie{u}(1)}(q,t_1,t_2,r)$ of the holomorphic twist of the theory on a single membrane is given by the plethystic exponential of the single particle index
\beqn\label{eqn:3done}
f^{3d}_{\lie{u}(1)} (t_1,t_2,t_3,q) = \frac{q^{1/4} (t_1+t_2+t_3+ (t_1 t_2 t_3)^{-1}) - q^{3/4} (t_1^{-1} + t_2^{-1} + t_3^{-1} + t_1 t_2 t_3)}{1-q} .
\eeqn
\end{prop}

This agrees with the membrane index computed in \cite[Equation (2.31)]{Bhattacharya:2008zy}.

\subsection{A conjectural description for the theory two membranes}

\begin{prop} \label{prop:3dtwo}
The character of local operators of the factorization algebra $\clie_\bu(\cG_{2,c})$ on $\C^3$ is given by the following expression
\begin{multline}\label{eqn:3dtwo1}
(1-q) f^{3d}_{\lie{sl}(2)}(t_1,t_2,t_3,q) = q^2 + q (\chi_{\C^4}(t) \chi_{\C^4}(t^{-1}) - 1) \\ - q^{3/2} \frac12 (\chi_{\C^4}(t)^2 - \chi_{\C^4}(t^2)) - q^{1/2} \frac12 (\chi_{\C^4}(t)^2 + \chi_{\C^4}(t^2))
\end{multline}
\end{prop}

The Higgs and Coloumb branch specializations are given by 
\beqn
t_1 t_2 = q^{\pm 1/2} .
\eeqn
This results in the following expression
\beqn
f^{3d}_{\lie{sl}(2)}|_{t_1 t_2 = q^{\pm 1/2}} = q^{-1/2} t_1^{-2} + 1 + q^{1/2} t_1^2 .
\eeqn

\subsection{The membrane decomposition}

We briefly highlight an analogous construction of the restricted factorization algebra describing membranes.
Recall that in the twist we are considering the worldvolume theory of membranes is a three-dimensional partially topological-holomorphic field theory which exist on a product $\R \times C$ where $C$ is a Riemann surface.

We consider the eleven-dimensional theory on $\R \times X_C$ where $X_C$ is the Calabi--Yau fivefold
\[
X_C = \text{Tot}(K_C^{1/4} \otimes \C^4),
\]
the total space of the rank four bundle $K_C^{1/4} \otimes \C^4$ over $C$.
Denote by 
\[
\pi \colon V \to \R \times C
\] 
the pullback of this vector bundle along the projection $\R \times C \to C$.
We can equivalently think of the eleven-manifold as the total space of the bundle $V$ over $\R \times C$. 
In this description, the twisted membranes will be supported along the zero section of $V$. 
The factorization algebra $\Obs_{sugra}$ on ${\rm Tot}(V)$ can be pushed forward along $\pi$ to a factorization algebra $\pi_* \Obs_{sugra}$ on $\R \times C$.

As in the fivebrane case, this is not the factorization algebra associated to an ordinary field theory on $\R \times C$.
Nevertheless there is a subfactorization algebra $\Bar{\pi}_* \Obs_{sugra}$ which admits a natural grading so that each filtered component can be understood as such.

To construct this factorization algebra we consider the action of the compact torus $U(1)^{\times 4}$ which rotates the fibers of the bundle $K_C^{1/4} \otimes \C^4$.
We let $\Bar{\pi}_* \Obs_{sugra}$ be the factorization algebra on $\R \times C$ consisting of observables which are finite sums of integral eigenvectors for this group action.
Again, we find that $\Bar{\pi}_* \Obs_{sugra}$ is of the form $\clie^\bu(\cG_{\R \times C})$ for some $\infty$ sheaf of $L_\infty$ algebras $\cG_{\R \times C}$ on $\R \times C$ which is the sections of a pro vector bundle.

Analogous to the decomposition determined by the $\C^\times$ action considered in \ref{s:cstarfive}, there is a decomposition of $\cG_{\R \times C}$ determined by a (different) $\C^\times$ acting on the fields of the eleven-dimensional theory on ${\rm Tot}(V) \cong \R \times X_C$.

In a local THF chart $U \subset \R \times C$ we can write a generic field of the eleven-dimensional theory on $\pi^{-1} U$ in coordinates as $f(t;z,w)$ where $(t;z) \in U \subset\R \times C$ is THF coordinate and $w=(w_1,w_2,w_3,w_4)$ is the holomorphic fiber coordinate of the rank four bundle $V$.

%For simplicity let us consider the eleven-dimensional theory on flat space $\R \times \C^5$ with some number of fivebranes supported on
%\[
%0 \times 0 \times \C^3 \subset \C^5 .
%\]
%is that (after taking into account the backreaction) the observables on a large number of fivebranes is Koszul dual to the factorization algebra 
%
%\parsec[s:flatdecomp]
%
%%In the simple case where $Z = \C^3$ and we identify the total space of $K^{1/2}_Z \otimes \C^2$ with $\C^5$ then the manifold obtained by removing the locus of the brane is homeomorphic to
%%\[
%%\C^3 \times (\R \times \C^2 - 0) .
%%\]
%%
%%Let $\pi \colon \R \times \C^5 - (0 \times \C^3) \to \C^3 \times \R_{>0}$ be the projection map whose fibers are homeomorphic to the sphere $S^4$ which links the location of the fivebranes.
%%We restrict the factorization algebra of the eleven dimensional theory $\Obs_{sugra}$ to the open set obtained by removing the locus of the brane.
%
%In the simple case that $Z = \C^3$ and we identify the total space of $K^{1/2}_Z \otimes \C^2$ with $\C^5$ there is a more direct construction of the factorization algebra $\Obs_{sugra}|_{Z}$. 
%
%Let $\pi \colon \R \times \C^5 \to \C^3$ be the projection map.
%Then, via $\pi$ we can pushforward the factorization algebra associated to the eleven-dimensional theory to obtain a factorization algebra
%\[
%\pi_* \Obs_{sugra} 
%\]
%on $\C^3$.
%This factorization algebra is not the factorization algebra associated to an ordinary sort of field theory on $\C^3$. 
%Nevertheless there is a subfactorization algebra which admits a natural grading so that each filtered component can be understood as such.

With this notation in place, we introduce the following $\C^\times$ action on the fields of the eleven-dimensional theory:
\begin{itemize}
\item On the fields $\mu(t;w,z)$ the action is
\[
\lambda \cdot \mu(t;z,w) = \mu(t; z , \lambda w).
\]
\item On the fields $\nu(t;w,z)$ the action is
\[
\lambda \cdot \nu(t;z,w) = \nu(t;z ,\lambda w).
\]
\item On the fields $\beta(t;w,z)$ the action is
\[
\lambda \cdot \beta(t;z,w) = \lambda^{-2} \beta(t; z, \lambda w).
\]
\item On the fields $\gamma(t;w,z)$ the action is
\[
\lambda \cdot \gamma(t;z,w) = \lambda^{-2} \gamma(t; z ,\lambda w).
\]
\end{itemize}

\begin{prop}
The $L_\infty$ structure on $\cL_{sugra}$ is equivariant for this $\C^\times$ action. 
%The odd symplectic form on $\cL_{sugra}$ is weight $-1$ for this $\C^\times$ action. 
\end{prop}

As in the case of fivebranes, this proposition allows us to decompose the local Lie algebra $\cG_{\R \times C}$ into weights. 
Similarly as above one can show that the $n$th eigenspace is quasi-isomorphic to a local Lie algebra of the form
\[
\Omega^\bu(\R) \otimes \Omega^{0,\bu}\left(C, \cV^{(n)}\right)
\]
where $\cV^{(n)}$ is a holomorphic super vector bundle on $C$.

We will state explicit descriptions of the first few eigenspaces without details of the proof. 
The weight decomposition is concentrated in degrees $\geq -2$. 
The lowest weight $n=-2$ is not very interesting, it is quasi-isomorphic just to the de Rham complex 
\beqn
\Omega^\bu (\R) \otimes \Omega^{\bu, \bu}(\C) \simeq \Omega^\bu (\R^3) .
\eeqn

The next nontrivial eigenspace is $n=-1$. 
Here, one can see that 
\[
\cV^{(-1)} \simeq K_{C}^{1/4} \otimes \C^4 \oplus \Pi K^{3/4}_C \otimes \C^4 .
\]
For $n=0$ one has 
\[
\cV^{(0)} \simeq \T_C \oplus \lie{sl}(4) \otimes \cO_C \oplus \Pi \left(K^{-1/2}_C \otimes \wedge^2 \C^4 \oplus K^{-1/2}_C \otimes S^2 \C^4\right) .
\]
As in the case of fivebranes, the $n=0$ eigenspace is equipped with the structure of a super Lie algebra whose exact form we do not state now.

\end{document}