### 2023-04-17 fivebranes/IIA.tex
IIA

Let $S$ be a complex surface.
Let $p_W \colon W \to S$ be a rank two holomorphic vector bundle on $S$ and let $p_L \colon L \to S$ be a holomorphic line bundle on $S$.
We require that these bundles satisfy
\beqn
\det (W \oplus L) = K_S 
\eeqn
where $K_S$ is the canonical bundle of $S$.
This condition guarauntees that
\beqn
X = \text{Tot}(W \oplus L)
\eeqn
is a Calabi--Yau five-fold.
We will also assume that $L$ has a real structure with compatible fiberwise hermitian metric.

We place the holomorphic twist of $M$-theory on 
\beqn
\R \times X
\eeqn
and consider fivebranes wrapping $Z = \text{Tot}(L)$.

Reduction to IIA is described by a diagram
\beqn
\begin{tikzcd}
S^1 \ar[r] \ar[d] & X \ar[d] & \ar[l] {pt} \ar[d] \\
(pt \ne 0, t, z) \ar[r] & \text{Tot}(\R_{\geq 0} \to \R \times \text{Tot}(W)) & \ar[l] (0,t,z) .
\end{tikzcd} .
\eeqn
When $L$ is trivial this is simply
\beqn
\R \times \C \to \text{Tot}(W) \to \R \times \R_+ \times \text{Tot}(W).
\eeqn
The fivebranes become fourbranes wrapping
\beqn
\text{Tot}(\R_+ \to \text{Tot}(W)) .
\eeqn
When $L$ is trivial then the fourbranes wrap
\beqn
0 \times \R_+ \times S .
\eeqn

Equations of motion for the open string theory describes moduli space of pairs $\{(\cL, \cF)\}$ where $\cL$ is a local system on $\R_+$ of rank one and $\cF$ is a perfect complex on $\text{Tot}(W)$ which is finite over $S$.
This space is naturally $(-3)$-shifted symplectic.
If we polarize $W = L_1 \oplus L_2$ to $L_1 \oplus L_2[2]$ then this becomes $(-1)$-shifted symplectic.
There is a forgetful functor to just remembering the perfect complex over $\text{Tot}(W)$ (which is fine over $S$), and the claim is that this is $0$-shifted symplectic.