### 2023-05-11 graviton spectra/main.tex

\documentclass{amsart}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\let\ref\hyperref
\let\hat\widehat
\usepackage{amsfonts} %necessary for Blackboard Bold
\usepackage{amsmath} %does A LOT
\usepackage{amssymb} %necessary for things like \partial
\usepackage{amsthm} %for thms and such
\usepackage{macros}
\usepackage{tikz-cd}
\usepackage{tikz-feynman}
\usepackage{xparse}
\usepackage[centerboxes,boxsize=1ex,nobaseline,aligntableaux=center]{ytableau}
\setcounter{secnumdepth}{4}

\title{Twisted graviton spectra of $AdS_4\times S^7$ and $AdS_7\times S^4$} 
\author{Surya Raghavendran and Brian R. Williams}
\begin{document}

\maketitle

\begin{abstract} %required
\end{abstract}

\tableofcontents

Having described our eleven-dimensional model on flat spacetimes, we now pursue descriptions on maximally symmetric spacetimes. We begin by describing twisted versions of the $AdS_4\times S^7$ and $AdS_7\times S^4$ backgrounds. In eleven-dimensional supergravity, these backgrounds arise as near-horizon limits of the backgrounds sourced by some number of M2 and M5 branes in flat space respectively. In the first section of this chapter \ref{sec:ads}, we describe an analogous procedure natively in our twisted context.

To do so, we motivate ansatzae for the leading order couplings of our eleven-dimensional model to M2 and M5 branes. These couplings determine certain curved deformations of the $L_\infty$-algebra underlying our model - we conjecture that deforming the theory in the complement of the brane by a solution to the resulting curved Maurer-Cartan equation is perturbatively equivalent to the twist of the theory on $AdS_4\times S^7$ and $AdS_7\times S^4$.

The next two sections provide evidence for this conjecture. We begin in section \ref{sec:states} with numerical checks. We give definitions of supergravity states in our twisted $AdS_4\times S^7$ and $AdS_7\times S^4$ backgrounds, which can be thought of as particular field configurations that are localized at points on the conformal boundary of $AdS$. We compute characters of the proposed state spaces and find exact matches with counts of gravitons on $AdS_4\times S^7$ and $AdS_7\times S^4$ respectively.

The next strand of evidence we pursue is by matching symmetries. In the physical theory, the $AdS_4\times S^7$ and $AdS_7\times S^4$ backgrounds carry actions of the 3d $\mc N=8$ and 6d $\mc N=(2,0)$ superconformal algebras. We show that our conjectural descriptions of twists of these backgrounds carry actions of the minimal twists of the corresponding superconformal algebras.

With these pieces of evidence in hand, in sections \ref{sec:e16}, \ref{sec:e36},  we then turn to study some representation theoretic aspects of the state spaces constructed in section \ref{sec:states}. We identify certain $\C^\times$ actions on our eleven-dimensional model that combine rescalings in directions normal to branes with a certain rescaling of the space of fields - this induces a decomposition of the space of fields that we dub the \textit{graviton decomposition}. The weight $0$ parts of these decompositions are certain local $L_\infty$-algebras whose costalks recover the linearly compact super-Lie algebras $E(1|6)$ and $E(3|6)$. We thus see that these linearly compact super-Lie algebras act on the spaces of supergravity states constructed in section \ref{sec:states}. We explicate their action on nonzero weight spaces of the graviton decomposition.

In the final section of the chapter, we motivate some current work in progress that leverages the uncovered appearance of exceptional linearly compact super-Lie algebras for holographic means. Eleven-dimensional supergravity on $AdS_4\times S^7$ and $AdS_7\times S^4$ is expected to be equivalent to the large $N$ limit of the worldvolume theories of $N$ M2 branes and $N$ M5 branes respectively. 
%Typically, this definition is made in situations where the relevant boundary value problem has a unique solution, in which case one may label states by the corresponding boundary values. Moreover, one may think of such boundary values as arising from modifications of a vacuum boundary condition at a point.

\input{content/br}

\input{content/index}

\input{content/suco}

\input{content/e16}

\input{content/e36}

%\input{chapters/4-ads/ops}

%\input{chapters/4-ads/finite}

\input{content/holography}

\bibliographystyle{amsalpha}
\bibliography{bibliography.bib}

\end{document}