### 2024-08-12 agt/survey.tex
agt

\documentclass[11pt]{amsart}

\usepackage{macros-master,amsaddr,physics,amssymb}

\addbibresource{refs}

\renewcommand{\div}{\del_{\Omega}}
\renewcommand{\PV}{\mathrm{PV}}
\renewcommand{\op}{\operatorname}

\setcounter{tocdepth}{1}

\begin{document}
\title{Notes on agt}
\maketitle

\section{The twisted brane construction}

We will use the fact that the minimal twist of $M$-theory exists on geometries of the form
\beqn
Z \times \R
\eeqn
where $Z$ is a Calabi--Yau fivefold.

We consider familiar extended objects, which in the twisted setting have the following flavor.
\begin{itemize}
\item[(M5)] Twisted fivebranes wrap an embedded threefold
\beqn
X \subset Z
\eeqn
with the property that $K_X \simeq \wedge^2 N_{X \subset Z}$ where $N_{X \subset Z}$ is the normal bundle to $X$ in $Z$.
\item[(M2)] Twisted membranes wrap the real line times an embedded curve $\Sigma \subset Z$:
\beqn
\Sigma \times \R \subset Z \times \R
\eeqn
with the property that $K_\Sigma \simeq \wedge^4 N_{\Sigma \subset Z}$.
\end{itemize}

Some special cases of such twisted geometries are worth addressing.

\begin{itemize}
\item[(1)] First, consider the following local fivefold geometry
\beqn
Z = \op{Tot} \begin{pmatrix} W \\ \downarrow \\ X \end{pmatrix} 
\eeqn
where $W$ is a rank two vector bundle on a threefold $X$, which satisfies $\wedge^2 W = K_X$.
The case $X = \C^3$ and $W = K^{1/2}_{\C^3} \otimes \C^2$.
The worldvolume theory on $\C^3$ in this geometry is the (minimal) holomorphic twist of six-dimensional $\cN=(2,0)$ supersymmetry.
More generally, if a threefold $X$ admits a square root of its canonical bundle $K^{1/2}_X$, then we can consider the geometry 
\beqn
Z = \op{Tot} \begin{pmatrix} K^{1/2}_X \otimes \C^2 \\ \downarrow \\ X \end{pmatrix} ,
\eeqn
with fivebranes wrapping the zero section $X \subset Z$.
\item[(2)]
Notice that in the above example there is a global $SU(2)$ symmetry.
We can break this to $U(1)$ by assuming that $W = L_1 \oplus L_2$ where $L_1,L_2$ are line bundles on $X$ satisfying $L_1 \otimes L_2 = K_X$.
Note that necessarily we have $L_2 = L^{-1} \otimes K_X$, $L=L_1$.
\item[(3)] Suppose that $Y$ is a Calabi--Yau threefold with an embedded holomorphic curve $C$, and $S$ is a Calabi--Yau surface.
Then $Z = Y \times S$ determines a twisted $M$-theory geometry with fivebranes that wrap 
\beqn
X = C \times S \subset Y \times S \times \R = Z \times \R .
\eeqn
\item[(4)]
As a special case of (3), consider the Calabi--Yau threefold
\beqn
Y = \op{Tot} \begin{pmatrix} \C \oplus \T^*_C \\ \downarrow \\ C \end{pmatrix} \simeq \C \times \T^* C ,
\eeqn
where $C$ is an arbitrary Riemann surface.
To phrase this as a special case of (2) one takes $W$ to be the rank two bundle obtained by restricting $\C \oplus \T^*_C$ along $C \times S \to C$.
In this geometry, one can allow for $C$ to have punctures.
The four-dimensional theory one obtains by compactifying the six-dimensional worldvolume theory of fivebranes along
\beqn
X = C \times S \to S 
\eeqn
in this geometry gives rise to the holomorphic twist of four-dimensional theories of class $\cS$ \cite{GaiottoS,GMN}.
\item[(5)] 
Again, as a special case of (3) (but more general than the last example), consider the Calabi--Yau threefold
\beqn
Y = \op{Tot} \begin{pmatrix} V \\ \downarrow \\ C \end{pmatrix} ,
\eeqn
where $V$ is an $SU(2)$ bundle over $C$.
Compactification along $X = C \times S \to S$ gives rise to holomorphic twists of four-dimensional superconformal theories \cite{BeemM5}.
This fits as a special case of (1) by taking $W$ to be the restriction of the bundle $V$ along the projection $X \to C$.
\end{itemize}

\section{Twisted fivebranes}

We will work in the generality of the geometry in (1) where the fivefold is of the form $\op{Tot}(W \to X)$ with $\wedge^2 W = K_X$.

The theory of a twisted fivebrane in this geometry admits the following description.
In the Poisson BV formalism its sheaf of fields is of the form $\Omega^{0,\bu}(X,\mathbb{V})$ where $\mathbb{V}$ is the following super complex of holomorphic vector bundles
\beqn
\mathbb{V} \;\;\;\left\{
\begin{tikzcd}
& \ul{-1} & \ul{0} \\
\ul{\text{even}} & \Omega^2_X \ar[r,"\del"] & \Omega^3_X \\
\ul{\text{odd}} & \Pi W  
\end{tikzcd} \right.
\eeqn
We will write Dolbeault valued fields of the bundle in the first line with the letter as~$\beta$ and in the bottom line as $\phi$.

The Poisson BV structure is determined by\brian{finish}

\subsection{Five-dimensional supersymmetric Yang--Mills theory}

Let $S$ be a complex surface and $W$ a rank two vector bundle on $S$ satisfying $\wedge^2 W = K_S$.
Given any Lie algebra~$\lie{g}$ equipped with a non-degenerate invariant pairing, we consider the following $\Z \times \Z/2$ graded five-dimensional gauge theory which we refer to as generalized twisted five-dimensional $\cN=2$ Yang--Mills theory.

The fields are
\begin{itemize}
\item Hybrid $BF$ theory with fields
\begin{align*}
A & \in \Omega^{\bu}(\R) \otimes \Omega^{0,\bu}(S) \otimes \lie{g} [1] \\
B & \in \Omega^{\bu}(\R) \otimes \Omega^{2,\bu}(S)\otimes \lie{g} [1] .
\end{align*}
\item Hypermultiplet fields
\beqn
\phi \in \Omega^\bu(\R) \otimes \Omega^{0,\bu}(S,\Pi W) [1] .
\eeqn
In other words, $\phi^{0,0}$ is an odd field of cohomological degree $-1$.
\end{itemize}
The action functional is
\beqn
\int_{\R \times S} (B , F_A) + \frac12 \int_{\R \times S} (\phi, \dbar_A \phi) .
\eeqn

Alternatively, this theory is hybrid Chern--Simons theory associated to the cyclic, holomorphic, local Lie superalgebra
\beqn
\lie{g} \otimes \wedge^\bu W ,
\eeqn
where $\wedge^0, \wedge^2$ are even and $\wedge^1$ is odd.
Here, we view $\wedge^\bu W$ as a bundle of commutative super algebras.
Explicitly, the nonzero brackets $[-,-]$ in this Lie superalgebra is of the form
\begin{align*}
\lie{g} \otimes \wedge^0 W \times \lie{g} \otimes \wedge^0 W & \to \lie{g} \otimes \wedge^0 W  \\
\lie{g} \otimes \wedge^0 W \times \lie{g} \otimes \wedge^1 W & \to \lie{g} \otimes \wedge^1 W \\
\lie{g} \otimes \wedge^1 W \times \lie{g} \otimes \wedge^1 W & \to \lie{g} \otimes \wedge^2 W .
\end{align*}

The perturbative phase space on $S$ is the $\Z \times \Z/2$ even symplectic space
\beqn
\T^* \left( \lie{g} \otimes \Omega^{0,\bu}(S)[1] \right) \oplus \Pi \left(\lie{g} \otimes \Omega^{0,\bu}(S,W)[1] \right) .
\eeqn
A moduli theoretic interpretation of this is as follows.
%Consider the moduli stack $\op{Bun}_G(S)$ of holomorphic $G$-bundles on $S$.
%Let $\T^*_{\op{Bun}_G(S)}$ be its cotangent complex.
Consider the moduli of triples
\beqn
(P, \eta, \phi)
\eeqn
where $P$ is a holomorphic $G$-bundle on $S$, $\eta \in \Omega^{2,1}(S, \op{ad} P)$, and $\phi \in \Omega^{0,1}(S,\op{ad}(P) \otimes W)$ which satisfy \brian{equations}.

Let us now assume we have a polarization of $W = L_1 \oplus L_2$ into a sum of two line bundles on $S$ satisfying $L_1 \otimes L_2 = K_S$.
Without loss of generality we set $L = L_1$ and $L_2 = L^{-1} \otimes K_S$.
This allows us to consider a tilted (regraded) perturbative phase space\footnote{We should explain $\Pi W[1] \to L_2[2] \oplus L_1$.}
\beqn
\begin{tikzcd}
\ul{-2} & \ul{-1} & \ul{0} \\
& \Omega^{0,\bu}(S, \lie{g}) & \\
\Omega^{2,\bu}(S,\lie{g} \otimes L^{-1}) & & \Omega^{0,\bu}(S, \lie{g} \otimes L) \\
& \Omega^{2,\bu}(S, \lie{g}) & .
\end{tikzcd}
\eeqn
Then, the perturbative phase space of generalized, twisted, five-dimensional $\cN=2$ supersymmetric Yang--Mills theory can be written as
\beqn
\op{Higgs}_G (S,L) .
\eeqn
Here $\op{Higgs}_G (S,L)$ is the moduli stack of quadruples $(P,\varphi, \varphi^*,B)$ where $P$ is a holomorphic $G$-bundle,
\begin{align*}
\varphi & \in \Gamma(S, \op{ad}(P) \otimes L) \\
\varphi^* & \in \Omega^{2,2}(S, \op{ad}(P) \otimes L^{-1}) \\
B & \in \Omega^{2,1}(S, \op{ad} P) 
\end{align*}
satisfying
\begin{align*}
\dbar \varphi = \dbar \varphi^* & = 0 \\
\dbar B + [\varphi, \varphi^*] & = 0 .
\end{align*}
We can furthermore write this as a \textit{twisted} cotangent bundle to the moduli stack $\cM_{G}(S,L)$ of pairs $(P,\varphi)$ where $P$ is a holomorphic principal $G$-bundle and $\varphi \in \Gamma(S, \lie{g} \otimes L)$ satisfying $\dbar \varphi = 0$.
The twist is by the two form \brian{finish}.

\section{AGT statements}

Let $S$ be a projective surface and $L_1,L_2,L_3$ line bundles on $S$ satisfying 
\beqn
L_1 \otimes L_2 \otimes L_3 = K_S .
\eeqn
We consider a special case of the geometry (2) where the threefold that twisted fivebranes $X$ is of the form
\beqn
X = \op{Tot} \begin{pmatrix} L_3 \\ \downarrow \\ S \end{pmatrix},
\eeqn
and where $W = L_1 \oplus L_2$ (or rather, the restrictions of these line bundles to $X$).
In this case, the local Calabi--Yau fivefold geometry is $Z = \op{Tot}(L_1 \oplus L_2 \oplus L_3).$
\end{document}