### 2024-08-14 agt/survey.tex
modes

\documentclass[11pt]{amsart}

\usepackage{macros-master,amsaddr,physics,amssymb}

\addbibresource{refs}

\renewcommand{\div}{\del_{\Omega}}
\renewcommand{\PV}{\mathrm{PV}}
\renewcommand{\op}{\operatorname}
\newcommand{\CB}{\mathbb{C}\mathbb{B}}
\newcommand{\SB}{\mathbb{S}\mathbb{B}}

\setcounter{tocdepth}{1}

\begin{document}
\title{Notes on agt}
\maketitle

\section{The twisted brane construction}

We will use the fact that the minimal twist of $M$-theory exists on geometries of the form
\beqn
Z \times \R
\eeqn
where $Z$ is a Calabi--Yau fivefold.

We consider familiar extended objects, which in the twisted setting have the following flavor.
\begin{itemize}
\item[(M5)] Twisted fivebranes wrap an embedded threefold
\beqn
X \subset Z
\eeqn
with the property that $K_X \simeq \wedge^2 N_{X \subset Z}$ where $N_{X \subset Z}$ is the normal bundle to $X$ in $Z$.
\item[(M2)] Twisted membranes wrap the real line times an embedded curve $\Sigma \subset Z$:
\beqn
\Sigma \times \R \subset Z \times \R
\eeqn
with the property that $K_\Sigma \simeq \wedge^4 N_{\Sigma \subset Z}$.
\end{itemize}

Some special cases of such twisted geometries are worth addressing.

\begin{itemize}
\item[(1)] First, consider the following local fivefold geometry
\beqn
Z = \op{Tot} \begin{pmatrix} W \\ \downarrow \\ X \end{pmatrix} 
\eeqn
where $W$ is a rank two vector bundle on a threefold $X$, which satisfies $\wedge^2 W = K_X$.
The case $X = \C^3$ and $W = K^{1/2}_{\C^3} \otimes \C^2$.
The worldvolume theory on $\C^3$ in this geometry is the (minimal) holomorphic twist of six-dimensional $\cN=(2,0)$ supersymmetry.
More generally, if a threefold $X$ admits a square root of its canonical bundle $K^{1/2}_X$, then we can consider the geometry 
\beqn
Z = \op{Tot} \begin{pmatrix} K^{1/2}_X \otimes \C^2 \\ \downarrow \\ X \end{pmatrix} ,
\eeqn
with fivebranes wrapping the zero section $X \subset Z$.
\item[(2)]
Notice that in the above example there is a global $SU(2)$ symmetry.
We can break this to $U(1)$ by assuming that $W = L_1 \oplus L_2$ where $L_1,L_2$ are line bundles on $X$ satisfying $L_1 \otimes L_2 = K_X$.
Note that necessarily we have $L_2 = L^{-1} \otimes K_X$, $L=L_1$.
\item[(3)] Suppose that $Y$ is a Calabi--Yau threefold with an embedded holomorphic curve $C$, and $S$ is a Calabi--Yau surface.
Then $Z = Y \times S$ determines a twisted $M$-theory geometry with fivebranes that wrap 
\beqn
X = C \times S \subset Y \times S \times \R = Z \times \R .
\eeqn
\item[(4)]
As a special case of (3), consider the Calabi--Yau threefold
\beqn
Y = \op{Tot} \begin{pmatrix} \C \oplus \T^*_C \\ \downarrow \\ C \end{pmatrix} \simeq \C \times \T^* C ,
\eeqn
where $C$ is an arbitrary Riemann surface.
To phrase this as a special case of (2) one takes $W$ to be the rank two bundle obtained by restricting $\C \oplus \T^*_C$ along $C \times S \to C$.
In this geometry, one can allow for $C$ to have punctures.
The four-dimensional theory one obtains by compactifying the six-dimensional worldvolume theory of fivebranes along
\beqn
X = C \times S \to S 
\eeqn
in this geometry gives rise to the holomorphic twist of four-dimensional theories of class $\cS$ \cite{GaiottoS,GMN}.
\item[(5)] 
Again, as a special case of (3) (but more general than the last example), consider the Calabi--Yau threefold
\beqn
Y = \op{Tot} \begin{pmatrix} V \\ \downarrow \\ C \end{pmatrix} ,
\eeqn
where $V$ is an $SU(2)$ bundle over $C$.
Compactification along $X = C \times S \to S$ gives rise to holomorphic twists of four-dimensional superconformal theories \cite{BeemM5}.
This fits as a special case of (1) by taking $W$ to be the restriction of the bundle $V$ along the projection $X \to C$.
\end{itemize}

\section{Twisted fivebranes}

We will work in the generality of the geometry in (1) where the fivefold is of the form $\op{Tot}(W \to X)$ with $\wedge^2 W = K_X$.

The theory of a twisted fivebrane in this geometry admits the following description.
In the Poisson BV formalism its sheaf of fields is of the form $\Omega^{0,\bu}(X,\bV)$ where $\bV$ is the following super complex of holomorphic vector bundles
\beqn\label{eqn:fivebranes}
\bV \;\;\;\left\{
\begin{tikzcd}
& \ul{-1} & \ul{0} \\
\ul{\text{even}} & \Omega^2_X \ar[r,"\del"] & \Omega^3_X \\
\ul{\text{odd}} & \Pi W  
\end{tikzcd} \right.
\eeqn
We will write Dolbeault valued fields of the bundle in the first line with the letter as~$\beta$ and in the bottom line as $\phi$.

The Poisson BV structure is determined by\brian{finish}

\subsection{Five-dimensional supersymmetric Yang--Mills theory}

Let $S$ be a complex surface and $W$ a rank two vector bundle on $S$ satisfying $\wedge^2 W = K_S$.
Given any Lie algebra~$\lie{g}$ equipped with a non-degenerate invariant pairing, we consider the following $\Z \times \Z/2$ graded five-dimensional gauge theory which we refer to as generalized twisted five-dimensional $\cN=2$ Yang--Mills theory.

The fields are
\begin{itemize}
\item Hybrid $BF$ theory with fields
\begin{align*}
A & \in \Omega^{\bu}(\R) \otimes \Omega^{0,\bu}(S) \otimes \lie{g} [1] \\
B & \in \Omega^{\bu}(\R) \otimes \Omega^{2,\bu}(S)\otimes \lie{g} [1] .
\end{align*}
\item Hypermultiplet fields
\beqn
\phi \in \Omega^\bu(\R) \otimes \Omega^{0,\bu}(S,\Pi W) \otimes \lie{g} [1] .
\eeqn
In other words, $\phi^{0,0}$ is an odd field of cohomological degree $-1$.
\end{itemize}
The action functional is
\beqn
\int_{\R \times S} (B , F_A) + \frac12 \int_{\R \times S} (\phi, \dbar_A \phi) .
\eeqn

Alternatively, this theory is hybrid Chern--Simons theory associated to the cyclic, holomorphic, local Lie superalgebra
\beqn
\lie{g} \otimes \wedge^\bu W ,
\eeqn
where $\wedge^0, \wedge^2$ are even and $\wedge^1$ is odd.
Here, we view $\wedge^\bu W$ as a bundle of commutative super algebras.
Explicitly, the nonzero brackets $[-,-]$ in this Lie superalgebra is of the form
\begin{align*}
\lie{g} \otimes \wedge^0 W \times \lie{g} \otimes \wedge^0 W & \to \lie{g} \otimes \wedge^0 W  \\
\lie{g} \otimes \wedge^0 W \times \lie{g} \otimes \wedge^1 W & \to \lie{g} \otimes \wedge^1 W \\
\lie{g} \otimes \wedge^1 W \times \lie{g} \otimes \wedge^1 W & \to \lie{g} \otimes \wedge^2 W .
\end{align*}

The perturbative phase space on $S$ is the $\Z \times \Z/2$ even symplectic space
\beqn
\T^* \left( \lie{g} \otimes \Omega^{0,\bu}(S)[1] \right) \oplus \Pi \left(\lie{g} \otimes \Omega^{0,\bu}(S,W)[1] \right) .
\eeqn
A moduli theoretic interpretation of this is as follows.
%Consider the moduli stack $\op{Bun}_G(S)$ of holomorphic $G$-bundles on $S$.
%Let $\T^*_{\op{Bun}_G(S)}$ be its cotangent complex.
Consider the moduli of triples
\beqn\label{eqn:pertphase1}
(P, B, \phi)
\eeqn
where $P$ is a holomorphic $G$-bundle on $S$, $\eta \in \Omega^{2,1}(S, \op{ad} P)$, and $\phi \in \Omega^{0,1}(S,\op{ad}(P) \otimes W)$ which satisfy 
\begin{align*}
\dbar \phi = 0 \\
\dbar B + \frac12 [\phi,\phi] = 0 .
\end{align*}
The perturbative phase space in \eqref{eqn:pertphase1} is the tangent complex to this moduli stack at the trivial $G$-bundle (and $B = \phi = 0$).

Let us now assume we have a polarization of $W = L_1 \oplus L_2$ into a sum of two line bundles on $S$ satisfying $L_1 \otimes L_2 = K_S$.
Without loss of generality we set $L = L_1$ and $L_2 = L^{-1} \otimes K_S$.
This allows us to consider a tilted (regraded) perturbative phase space\footnote{We should explain $\Pi W[1] \to L_2[2] \oplus L_1$.}
\beqn
\begin{tikzcd}
\ul{-2} & \ul{-1} & \ul{0} \\
& \Omega^{0,\bu}(S, \lie{g}) & \\
\Omega^{2,\bu}(S,\lie{g} \otimes L^{-1}) & & \Omega^{0,\bu}(S, \lie{g} \otimes L) \\
& \Omega^{2,\bu}(S, \lie{g}) & .
\end{tikzcd}
\eeqn
Then, the perturbative phase space of generalized, twisted, five-dimensional $\cN=2$ supersymmetric Yang--Mills theory can be written as the moduli stack
\beqn
\op{Higgs}_G (S,L) .
\eeqn
of quadruples $(P,\varphi, \varphi^*,B)$ where $P$ is a holomorphic $G$-bundle, and
\begin{align*}
\varphi & \in \Gamma(S, \op{ad}(P) \otimes L) \\
\varphi^* & \in \Omega^{2,2}(S, \op{ad}(P) \otimes L^{-1}) \\
B & \in \Omega^{2,1}(S, \op{ad} P) 
\end{align*}
which satisfy
\begin{align*}
\dbar \varphi = \dbar \varphi^* & = 0 \\
\dbar B + [\varphi, \varphi^*] & = 0 .
\end{align*}
We can furthermore write this as the cotangent bundle to the moduli stack $\cM_{G}(S,L)$ of pairs $(P,\varphi)$ where $P$ is a holomorphic principal $G$-bundle on $S$ and $\varphi \in \Gamma(S, \lie{g} \otimes L)$ satisfying $\dbar \varphi = 0$.

In this example, the Hilbert space of states coming from geometric quantization is functions on the base of this cotangent bundle
\beqn
\cO \left( \cM_{G}(S,L) \right) .
\eeqn 
Notice that when $L = K_S$ we have the Hilbert space
\beqn
\cO \left( \T[1] \op{Bun}_G(S) \right) \simeq \Omega^{\bu}(\op{Bun}_G(S)) .
\eeqn

\subsection{Chern--Simons deformation}

Let $\omega \in \Omega^{1,1}(S)$ be a $(1,1)$-form on $S$ with the property that $\dbar \omega = 0$ and $\del \omega = 0$.
In other words, $\omega$ defines an element in $H^1(S, \Omega^{1,cl})$.
Then, consider the deformation of generalized twisted five-dimensional $\cN=2$ Yang--Mills theory defined by
\beqn
S_\theta = \frac12 \int_{\R \times S} \omega \wedge (A, \del A) .
\eeqn
Locally, we can find a primitive $\omega = \del \theta$ and this deformation takes the more symmetric form $\frac12 \int \theta \wedge (\del A, \del A)$.

Turning on this deformation leads to the deformed perturbative phase space
\beqn
\begin{tikzcd}
\ul{-2} & \ul{-1} & \ul{0} \\
& \Omega^{0,\bu}(S, \lie{g}) \ar[dd,dotted,"\omega \wedge \del"] & \\
\Omega^{2,\bu}(S,\lie{g} \otimes L^{-1}) & & \Omega^{0,\bu}(S, \lie{g} \otimes L) \\
& \Omega^{2,\bu}(S, \lie{g}) & .
\end{tikzcd}
\eeqn
A more invariant way to construct this perturbative phase space is the following.
The non-degenerate invariant pairing determines a four-form $\kappa$ on $BG$.
This pulls back, along the evaluation map
\beqn
\op{ev} \colon S \times \op{Bun}_G(S) \to BG 
\eeqn
to a four-form on $S \times \op{Bun}_G(S)$.
Integration of the six-form $\omega \wedge \op{ev}^*\kappa$ along $S$ yields a two-form $\omega_{\kappa} = \int_S \omega \wedge \op{ev}^*\kappa$ on $\op{Bun}_G(S)$.
We can, in turn, further restrict this to a two-form on $\cM_{G}(S,L)$ that we will abusively still denote by $\omega_{\kappa}$.
The phase space is then the twisted cotangent bundle

\beqn\label{eqn:5dphase}
\T_{\omega_{\kappa}}^* \left(\cM_{G}(S,L) \right) .
\eeqn

\section{The abelian case}

We will work in the generality of the geometry in (1) where the fivefold is of the form $\op{Tot}(W \to X)$ with $\wedge^2 W = K_X$.

\subsection{Nakajima, Grojnowksi, and the Heisenberg vertex algebra}

Assume that $S$ is a projective surface equipped with a rank two vector bundle $V$ satisfying $\wedge^2 W = K_S$.
Consider the resulting twisted fivebrane worldvolume theory on $X = S \times C$ where $W = V \boxtimes K_C^{1/2}$ (here, we are assuming that $C$ is equipped with a spin structure).
In the Poisson BV formalism its sheaf of fields is of the form $\Omega^{0,\bu}(S \times C,\bV)$ where $\bV$ is the following super complex of holomorphic vector bundles
\beqn
\bV \colon \;\;\;\left\{
\begin{tikzcd}
& \ul{-1} & \ul{0} \\
\ul{\text{even}} & \Omega^2_{S \times C} \ar[r,"\del"] & \Omega^3_{S \times C} \\
\ul{\text{odd}} & \Pi p^* W  
\end{tikzcd} \right.
\eeqn
We will write Dolbeault valued fields of the bundle in the first line with the letter as~$\beta$ and in the bottom line as $\phi$, as we did above.
The BV quantization of this theory yields a factorization algebra on $S \times C$ that we will denote $\Obs_{S,W}$.
Its compactification, or factorization homology, along $S$
\beqn
\int_S \Obs_{S,W} 
\eeqn
is a holomorphic factorization algebra on $C$.
On $C = \C$ this determines a vertex algebra that we denote $\VV_{S,W}$.

We describe this vertex algebra explicitly.
Notice that since $S$ is projective there is a quasi-isomorphism of sheaves on $C$:
\begin{align*}
\Omega^{2,\bu}_{S \times C} & \simeq H^{\bu} (S, K_S) \otimes \Omega^{0,\bu}_C  \oplus H^\bu(S, \Omega^1_S) \otimes \Omega^{1,\bu}_C \\
\Omega^{3,\bu}_{S \times C} & \simeq H^{\bu} (S, K_S) \otimes \Omega^{1,\bu}_C  \\
\Omega^{0,\bu}_{S \times C}(W \boxtimes K_{C}^{1/2}) & \simeq H^\bu(S, W) \otimes \Omega^{0,\bu}_C(K^{1/2}_C) .
\end{align*}

Suppose that $V$ is a graded vector space equipped with a graded symmetric pairing of total degree zero.
Let $\{e_i\}$ be a basis for $V$.
There is a vertex algebra $\CB[V]$ generated by fields $e_i(z)$ of spin one and with OPEs
\beqn
e_i(z) e_j(w) \simeq \frac{(e_i,e_j)}{(z-w)^2} .
\eeqn
This is a graded vertex algebra with grading inherited from that of $V$.
Notice when $V = \C$ with the obvious pairing this is simply the chiral boson vertex algebra.

Similarly, suppose that $U$ is a $\Z \times \Z/2$ vector space equipped with a graded totally anti-symmetric pairing $\<-,-\>$ of total degree zero.
Then we can consider the symplectic boson vertex algebra valued in $U$ that we denote $\SB[U]$.
If $\{f_i\}$ is a basis for $U$ then the vertex algebra is generated by free fields $f_i(z)$ of spin $1/2$ and with OPEs
\beqn
f_i(z) f_j(w) \simeq \frac{\<f_i,f_j\>}{z-w} .
\eeqn
This vertex algebra inherits a $\Z \times \Z/2$ grading from $U$.

\begin{prop}
The vertex algebra $\VV_{S,W}$ is quasi-isomorphic to the vertex algebra
\beqn
\CB[H^\bu(S,\Omega^1_S)] \otimes \SB[\Pi H^\bu(S,W)[1]] \otimes T
\eeqn 
where $T$ is the vector space
\beqn
\Sym(H^\bu(S,K_S)[1])
\eeqn
equipped with the trivial vertex algebra structure.
\end{prop}

We now focus on a different, more familiar, choice of bundle data.
Suppose $S$ is a projective surface as before.
This time, we choose 
\beqn
W = \cO_S \boxtimes K_C \oplus K_S \boxtimes \cO_C .
\eeqn
and further regrade using the fact that $SU(2)$ has been broken to $U(1)$.
This six-dimensional theory is of the Dolbeault complex of the $\Z$-graded complex of holomorphic vector bundles
\beqn
\begin{tikzcd}
\ul{-2} & \ul{-1} & \ul{0} \\
& \Omega^2_{S \times C} \ar[r,"\del"] & \Omega^3_{S \times C} \\
\cO_{S} \boxtimes \Omega^1_C & & \Omega^2_S \boxtimes \cO_C .
\end{tikzcd} 
\eeqn
The BV quantization of this theory yields a factorization algebra on $S \times C$ that we will denote $\Obs_{S}^{NG}$.
Its compactification, or factorization homology, along $S$
\beqn
\int_S \Obs^{NG}_{S}
\eeqn
is a holomorphic factorization algebra on $C$.
On $C = \C$ this determines a vertex algebra that we denote $\VV_{S}^{NG}$.

\begin{prop}
The vertex algebra $\VV_{S}^{NG}$ is quasi-isomorphic to
\beqn
\CB[H^\bu(S,\Omega^1_S)] \otimes \beta\gamma[H^\bu(S,K)] \otimes T
\eeqn 
where $T$ is the vector space
\beqn
\Sym(H^\bu(S,K_S)[1])
\eeqn
equipped with the trivial vertex algebra structure.
The character of this vertex algebra (whose fugacities we explain in the proof) recovers G\"ottsche's formula for the Hodge-Poincar\'e polynomial of the Hilbert schemes of $S$:
\beqn
\chi(\VV_S^{NG}) = \prod_{m \geq 1} \prod_{k,\ell =0}^2 \left(1 - (-1)^{k+\ell} x^{k+m-1} y^{\ell + m - 1} q^m \right)^{(-1)^{k+\ell+1} h^{k,\ell}(S)} .
\eeqn
\end{prop}
\begin{proof}
Let $\til q$ be the fugacity associated to the conformal weight operator $L_0$.
Let $u$ be the fugacity for cohomological degree.
Let $v$ be the `flavor' fugacity, where $\gamma$ is of weight one and $\beta$ is opposite.
Finally, let $w$ be a fugacity corresponding to the rather trivial symmetry which scales the generators of the commutative part of the vertex algebra with weight one.

Then the single particle characters of each of the fields is computed directly as follows (here $h^{k,\ell} = h^{k,\ell}(S)$).
\begin{itemize}
\item The $\gamma$ field in the $\beta\gamma$ system valued contributes
\beqn
\frac{v}{1-\til q} \left(h^{2,0} - u h^{2,1} + u^2 h^{2,2}\right) .
\eeqn
\item The $\beta$ field in the $\beta\gamma$ system valued contributes
\beqn
\frac{v^{-1} \til q}{1- \til q} \left(u^{-2} h^{0,0} - u^{-1} h^{0,1} + h^{0,2}\right) .
\eeqn
\item The chiral boson field contributes
\beqn
\frac{\til q}{1-\til q} \left(-u^{-1} h^{1,0} + h^{1,1} - u h^{1,2} \right) .
\eeqn
\item The commutative part of the vertex algebra contributes
\beqn
w \left(-u^{-1} h^{2,0} + h^{2,1} - u h^{2,2}\right) .
\eeqn
\end{itemize}
To recover G\"ottsche's formula we specialize
\beqn
\til q = xy q, \quad u = y, \quad v = x y^{-1}, \quad w = x .
\eeqn
\end{proof}

\section{AGT statements}

Let $S$ be a projective surface and $L_1,L_2,L_3$ line bundles on $S$ satisfying 
\beqn
L_1 \otimes L_2 \otimes L_3 = K_S .
\eeqn
We consider a special case of the geometry (2) where the threefold that twisted fivebranes $X$ is of the form
\beqn
X = \op{Tot} \begin{pmatrix} L_3 \\ \downarrow \\ S \end{pmatrix},
\eeqn
and where $W = L_1 \oplus L_2$ (or rather, the restrictions of these line bundles to $X$).
In this case, the local Calabi--Yau fivefold geometry is $Z = \op{Tot}(L_1 \oplus L_2 \oplus L_3).$

We will consider a variant of the factorization algebra $\Obs_{S, W}$ in the case where $L_3$ is possibly a nontrivial line bundle. Indeed, BV quantization of the fields \eqref{eqn:fivebrane} of a twisted fivebrane on $X$ yields a holomorphic factorization algebra $\Obs_{X, W}$ that specializes to $\Obs_{S, W}$ in the case where $L_3$ is the trivial line bundle on $S$ \surya{introduce earlier}

From the factorization algebra $\Obs_{X, W}$, we will produce a pair of an associative algebra and a module. To do so, we choose a Hermitian metric on $L_3$. Doing so gives us a norm map

\beqn
X\to X\times_S X \to S\times \C
\eeqn

which factors through the trivial $\R_{\geq 0}$-bundle on $S$ since a Hermitian metric is in particular positive definite. Let $p$ denote the composition of this map with the projection onto $\R_{\geq 0}$.

The geometry of the situation can be summarized in the following diagram

\[
\begin{tikzcd}
X\setminus S \ar[d,"\mathring p"'] \ar[r,hook] & X \ar[d,"p"] & \ar[l,hook',"0"'] S \ar[d] \\
\R_{>0} \ar[r,hook] & \R_{\geq 0} & \ar[l,hook'] \{0\}.
\end{tikzcd}
\]

The pushforward
\beqn
p_* \Obs_{X,W}
\eeqn

is therefore a factorization algebra on $\R_{\geq 0}$.

We may extract from this data, an associative algebra and a module as follows.

\begin{itemize}
\item Let $(a,b)\subset \R_{\geq 0}$. The sections $(p_* \Obs_{X, W})(a,b)$ are an associative algebra, which we will denote as $\AA(X, W)$.

This associative algebra can equivalently be computed by restricting away from $S$, and pushing forward along $\mathring p$. The fiber of this map at $r \in \R_{>0}$ is a copy of $S_r(L_3)\subset X$, the total space of the radius $r$ circle bundle in $L_3$. Notice that in the case when $L_3$ is trivial, this associative algebra is exactly the modes algebra of the vertex algebra $\VV_{S,W}$.

\item The costalk at zero $(p_*\Obs{X,W})_0$ is a module, which we denote as $\MM (X, W)$. 
\end{itemize}

The AGT relation asserts that $\MM(X,W)$ may equivalently be computed as the geometric quantization of \eqref{eqn:5dphase}

\end{document}